---
title: "Plots; Hardness, density and 3D data analysis"
author: Eduardo Paixão, Tegenu Gossa, Walter Gneisinger, João Marreiros, Sören Tholen,
  Ivan Calandra, Erella Hovers
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

---

# Content
This script reads and plots the data from the 3D volume loss and Leeb Rebound hardness. 3D volume loss was calculated using CloudCompare cloud-to-mesh distance tool. Hardness was measured on each rock using the device Equotip Leeb Impact device C (HLC).
For details on the methods and data acquisition, please visit the Materials and Methods section of the paper. 

The knit directory for this script is the project directory.

---

# Load packages
```{r}

library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(doBy)
library(ggrepel)
library(flextable)
library(readr)
library(ggpubr)

```

---

# Import and preview data
```{r}

hlcdata <- read_csv2("../rawdata/hlcdata.csv")
volumelossdata2 <- read_csv2("../rawdata/volumelossdata.csv")
densitydata <- read_csv2("../rawdata/densitydata.csv")

str(hlcdata)
str(volumelossdata2)
str(densitydata)

```

# Manipulate hardness data
```{r}

# HLC data, Multiple variables are stored in column names, so transform it into observations per row
hlcdatalong <- hlcdata %>% 
  pivot_longer(c(`reading1`, `reading2`,`reading3`, `reading4`, `reading5`, `reading6`, `reading7`, `reading8`, `reading9`, `reading10`), names_to = "hlcvalues", values_to = "Hardness")

write_csv(hlcdatalong, "../deriveddata/hlcdatalong.csv")

```

# Summarise data
```{r}

# function
nminmaxmeanmedsd <- function(x){
    y <- x[!is.na(x)]
    n_test <- length(y)
    min_test <- min(y)
    max_test <- max(y)
    mean_test <- mean(y)
    med_test <- median(y)
    sd_test <- sd(y)
    out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test)
    names(out) <- c("n", "min", "max", "mean", "median", "sd")
    return(out)
}

# computation

# hlc data
num.var <- 22:length(hlcdatalong)
hlc_stats <- summaryBy(.~Material + cubeid,  data=hlcdatalong [c("Material", "cubeid", names(hlcdatalong)[num.var])], FUN=nminmaxmeanmedsd)

# volume loss
num.var2 <- 5:length(volumelossdata2)
volumeloss_stats <- summaryBy(.~Material + Cycle +  Sample + cubeid, data=volumelossdata2[c("Material", "Cycle", "Sample", "cubeid", names(volumelossdata2)[num.var2])], FUN=nminmaxmeanmedsd)

# View results
hlc_stats
volumeloss_stats

# save the results 
write_csv(hlc_stats, "../stats/stats_hlc.csv")
write_csv(volumeloss_stats, "../stats/stats_volumeloss.csv")

# join data tables

corr_data <- full_join(hlc_stats, densitydata,
              by = join_by("cubeid", "Material"), keep = F)

corr_data2 <- full_join(corr_data, volumeloss_stats,
              by = join_by("cubeid", "Material"))

# save single dataset
write_csv(corr_data2, "../deriveddata/full_data.csv")

```

# Plot hardness data, organised by raw material
```{r}

hlcplot <- ggplot(hlcdatalong, aes(Material, Hardness, colour = Material)) +
  theme_classic() + 
  theme(legend.title = element_blank()) +
  geom_boxplot() +
  geom_jitter() +
  guides(color = FALSE)  +
  labs(y = "Hardness mean (HLC)", x = "Raw material", colour = "Raw Material")

ggsave("../plots/hlcdata.png")

print(hlcplot)

```

# Plot volume loss data, organised by raw material
```{r}

# boxplot with all values, shows dispersion

volumeplot <- ggplot(volumeloss_stats, aes(Material, Value.max, colour = Material)) +
  geom_boxplot() +
  guides(color = FALSE)  +
  labs(y = "Damage (absolute Hausdorff distance)", x = "Raw material", colour = "Raw Material")

ggsave("../plots/damage_rawmaterial.png")

print(volumeplot)

# scatterplot with mean values

scat_volumeplot <- ggplot(corr_data2, aes(Material, Value.max, colour = Material)) +
  geom_point() +
  guides(color = FALSE)  +
  labs(y = "Damage (absolute Hausdorff distance)", x = "Raw material", colour = "Raw Material")

ggsave("../plots/scat_damage_rawmaterial.png")

print(scat_volumeplot)

arranged <- ggarrange(volumeplot, scat_volumeplot, common.legend = TRUE, ncol = 2, nrow = 1)

print(arranged)

ggsave("../plots/arranged.png")

```

# Correlation between hardness, density, and volume loss

```{r}

# scatterplot to show relationship between variables (dependent, volume loss; and independent, density and hardness)

scat <- ggplot(corr_data2,aes(Hardness.mean, Value.max, color = Material)) + 
  geom_point() +
  labs(y = "Damage (absolute Hausdorff distance)", x = "Hardness mean (HLC)", colour = "Raw Material")

ggsave("../plots/hardvolume.png")

print(scat)

scat2 <- ggplot(corr_data2,aes(density, Value.max, color = Material)) + 
  geom_point() +
  labs(y = "Damage (absolute Hausdorff distance)", x = "Density", colour = "Raw Material")

ggsave("../plots/densityvolume.png")

print(scat2)

scat3 <- ggplot(corr_data2,aes(density, Hardness.mean, color = Material)) + 
  geom_point()+
  labs(y = "Hardness mean (HLC)", x = "Density", colour = "Raw Material")

ggsave("../plots/densityhard.png")

print(scat3)

arranged2 <- ggarrange(scat, scat2, scat3, common.legend = TRUE, ncol = 3, nrow = 1)

print(arranged2)

ggsave("../plots/arranged2.png")

```

# Anova analysis
```{r}
# one-way anova test the effect of raw material on volume loss

# The null hypothesis (H0) of the ANOVA is no difference (between raw materials) in terms of volume loss, and the alternative hypothesis (Ha) is that volume loss are different from one another (between raw materials).

# check if there are differences among group means
onewaymodel <- aov(Value.max ~ Material, data = corr_data2)
summary(onewaymodel)

# The p-value of the raw material variable is low (p < 0.001), so it appears that the type of raw material (at least one) has a real impact on volume loss.

# check normality
hist(onewaymodel$residuals)

library(car)
qqPlot(onewaymodel$residuals)

# check if model fits the assumption, the homogeneity of variances
par(mfrow=c(2,2))
plot(onewaymodel)
par(mfrow=c(1,1))

# The model fits the assumption of homoscedasticity

# ANOVA tells us if there are differences among group means (volume loss), but not where those differences are. To find out which groups are statistically different from one another, you perform a Tukey’s Honestly Significant Difference Tukey’s HSD post-hoc test for pairwise comparisons.

# check which groups are statistically different from one another
TukeyHSD(onewaymodel, conf.level=.95) 
plot(TukeyHSD(onewaymodel, conf.level=.95), las = 1, cex.axis=0.5)

# Some observations from the post-hoc test results:
# All raw materials are significantly different from each other. All p-value (here represented as p adj) are smaller than 0.05.

```

# Bootstrapping

## Bootstrapped Mann-Whitney U p-value (95% confidence interval) for volume loss (damage) by raw material.

```{r, echo=FALSE}

p_boot <- vector()

db_shorten <- corr_data2 %>% group_by(Sample) %>% summarise(Material = unique(Material), Volumeloss = mean(Value.max, na.rm=T )) %>% as.data.frame()

for (i in 1:1000) {
  
  set.seed(i)

  basalt <- db_shorten[db_shorten$Material == 'basalt',]
  scoria <- db_shorten[db_shorten$Material == 'scoria',]
  ignimbrite <- db_shorten[db_shorten$Material == 'ignimbrite',] 
  glassyignimbrite <- db_shorten[db_shorten$Material == 'glassyignimbrite',]
  pumaceousignimbrite <- db_shorten[db_shorten$Material == 'pumaceousignimbrite',]
 
  basalt_sample <- basalt[sample(1:nrow(basalt), nrow(basalt), replace=TRUE),]
  scoria_sample <- scoria[sample(1:nrow(scoria), nrow(scoria), replace=TRUE),]  
  ignimbrite_sample <- ignimbrite[sample(1:nrow(ignimbrite), nrow(ignimbrite), replace=TRUE),]
  glassyignimbrite_sample <- glassyignimbrite[sample(1:nrow(glassyignimbrite), nrow(glassyignimbrite), replace=TRUE),]   
  pumaceousignimbrite_sample <- pumaceousignimbrite[sample(1:nrow(pumaceousignimbrite), nrow(pumaceousignimbrite), replace=TRUE),]
  
  combined3 <- rbind(basalt_sample, scoria_sample, pumaceousignimbrite_sample, glassyignimbrite_sample, ignimbrite_sample)
  
  test_res <- kruskal.test(Volumeloss ~ Material, data = combined3)
  
  p_boot[i] <- round(test_res$p.value, 4)
  
} 

return(quantile(p_boot, c(0.025, 0.975)))


```


# Correlations
## Kendall rank correlation between hardness and volume loss

```{r, echo=FALSE,}

colnames(db_shorten) <- c("ID", "Material", "HLC", "Density", "Volume loss")

cor.test(db_shorten$HLC, db_shorten$`Volume loss`, method='kendall')

```


## Kendall rank correlation between density and volume loss

```{r, echo=FALSE,}

colnames(db_shorten) <- c("ID", "Material", "HLC", "Density", "Volume loss")


cor.test(db_shorten$Density, db_shorten$`Volume loss`, method='kendall')
```

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

---

END OF SCRIPT
